---
title: "HW6 Solution"
author: "Zhengwei Song"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 2

## Importing the dataset
```{r, warning = FALSE, message=FALSE}
homicide_raw <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

#### Creating a `city_state` variable, binary variable `resolution` indicating case disposition, omitting cities excluded victim race and Tulsa, AL with data entry mistake. Also, limiting `victim_race` is white or black, making `victim_age` numeric.
```{r, warning = FALSE, message=FALSE}
homicide_df = homicide_raw %>% 
    janitor::clean_names() %>%
    mutate(
        reported_date = as.Date(as.character(reported_date), format = "%Y%m%d"),
        city_state = str_c(city, state, sep = ", "),
        resolved = as.numeric(disposition == "Closed by arrest"),
        victim_age = as.numeric(victim_age),
        victim_race = fct_relevel(victim_race, "White")
    ) %>%
    relocate(city_state) %>%
    filter(city_state != c("Tulsa, AL","Dallas, TX","Phoenix, AZ","Kansas City, MO"),
           victim_race == c("White","Black"))
```

#### Applying `glm` to fit a logistic regression with resolved vs unresolved as the outcome; victim age, sex and race as predictors. Also, saving the output of `glm` as an R object.
```{r, warning = FALSE, message=FALSE}
baltimore_logistic = homicide_df %>%
    filter(city_state == "Baltimore, MD") %>%
    glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 
```

#### Applying the `broom::tidy` to the above object; Obtaining the estimate and confidence interval of the adjusted odds ratio, for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r, warning = FALSE, message=FALSE}
baltimore_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```
* Keeping all other variables fixed, in Baltimore, MD, homicides in which the victim is male are significantly less like to be resolved than those in which the victim is female.

## Running `glm` for each of the cities in the dataset; Extracting the adjusted odds ratio and CI, for solving homicides comparing male victims to female victims; 
```{r, warning = FALSE, message=FALSE}
allcities_logistic = homicide_df %>%
    glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 
```

```{r, warning = FALSE, message=FALSE}
allcities_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```
* Keeping all other variables fixed, in XXX,XXX,XXX, homicides in which the victim is male are significantly less like to be resolved than those in which the victim is female.

```{r, warning = FALSE, message=FALSE}

```
